### 2012-03-05 docs/cqm_practical/cqm_practical.tex
added latex template

\documentclass{article}

\input{preamble.tex}
\input{tikzstyles.tex}

\begin{document}

test

\end{document}