### 2012-03-05 docs/cqm_practical/tikzstyles.tex

\tikzstyle{every picture}=[baseline=-0.25em]
\tikzstyle{dotpic}=[scale=0.6]

\tikzstyle{dot}=[inner sep=0.7mm,minimum width=0pt,minimum height=0pt,fill=black,draw=black,shape=circle]
\tikzstyle{white dot}=[dot,fill=white]
\tikzstyle{red dot}=[dot,fill=red,font=\footnotesize\color{white}]
\tikzstyle{green dot}=[dot,fill=green,font=\footnotesize]
\tikzstyle{hadamard}=[rectangle,fill=yellow,draw=black,font=\footnotesize,inner sep=2pt]

\tikzstyle{diredge}=[->]